Synthesizing Quotient Lenses

### 2018-07-06 papers/QuotientLenses/main.tex
Added acknowledgement
A bit of polishing

These unambiguity constraints can be a little fiddly in practice. As a simple example of this, consider writing a regular expression for comma-separated lists of strings. Our first impulse might be to write it as,
\begin{lstlisting}
CSL = anychar+ . ("," . anychar+)*
\end{lstlisting}
(where . is concatenation), but this regular expression is ambiguous, as \lstinline|anychar| is a big union of a of single characters including comma.

While the unambiguity constraints consequently appear to compromise compositionality of QREs, they can usually be circumvented by making a small changes to the offending regular expression: for instance, in the preceding example, then we need to write,
\begin{lstlisting}
CSL = anycharexceptcomma+ . ("," . anycharexceptcomma+)*
\end{lstlisting}
(assuming \lstinline|anycharexceptcomma| denotes a big union of every single character string except comma).

@@

To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
from the benchmark suite of the original Optician system. These
benchmarks are a combination of custom benchmarks, benchmarks derived
from FlashFill~\cite{flashfill}, and benchmarks derived from
Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).  In these 39 benchmarks, 10 of
the data formats had to be modified to work with the bijectivity
constraints that Optician required. For instance, when one
representation permits whitespace where the other does not, we
modified the original version of the benchmark to allow more whitespace,
thereby restoring bijectivity (but altering the data format). With the
new QRE support, we were able to remove these alterations.  This
experience alone suggests that QREs make the lens development process
more flexible.

@@

A newer language which supports quotenting is FliPpr \cite{matsuda2013flippr}. FliPpr is a program transformation system that uses program inversion to produce a CFG parser from a pretty-printer. FliPpr includes biased choice combinators that choose canonical representatives for sets of strings. For example, one can define variants of (white)spaces with the choice operator as,
\begin{center}
\begin{lstlisting}
nil = text "" <+ space
space = (text " " <+ text "\n") <> nil
\end{lstlisting}
\end{center}
where \lstinline|nil| and \lstinline|space| pretty-print \lstinline|""| and \lstinline|" "| respectively, but represent zero-or-more and one-or-more whitespaces in parsing. FliPpr also allows annotations for optional parentheses in programs. These features enable a user to easily define a pretty-printer that also recognizes ugly strings that get mapped to a canonical representative modulo whitespace and parentheses quotienting.

@@

>>>>

These unambiguity constraints can be a little fiddly in practice. As a simple example of this, consider writing a regular expression for comma-separated lists of strings. Our first impulse might be to write it as,
\begin{lstlisting}
CSL = anychar+ . ("," . anychar+)*
\end{lstlisting}
(where . is concatenation), but this regular expression is ambiguous, as \lstinline|anychar| is a big union of a of single characters including comma.
Instead, in this case, we need to write
\begin{lstlisting}
CSL = anycharexceptcomma+ . ("," . anycharexceptcomma+)*
\end{lstlisting}
(assuming \lstinline|anycharexceptcomma| denotes a big union of every single character string except comma).

@@

To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
from the benchmark suite of the original Optician system. These
benchmarks are a combination of custom benchmarks, benchmarks derived
from FlashFill~\cite{flashfill}, and benchmarks derived from
Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparison data, and crime index data).  In these 39 benchmarks, 10 of
the data formats had to be modified to work with the bijectivity
constraints that Optician required. For instance, when one
representation permits whitespace where the other does not, we
modified the original version of the benchmark to allow more whitespace,
thereby restoring bijectivity (but altering the data format). With the
new QRE support, we were able to remove these alterations.  This
experience alone suggests that QREs make the lens development process
more flexible.

@@

A newer language which supports quotenting is FliPpr \cite{matsuda2013flippr}. FliPpr is a program transformation system that uses program inversion to produce a CFG parser from a pretty-printer. FliPpr includes biased choice combinators that choose canonical representatives for sets of strings. For example, one can define variants of (white)spaces with the choice operator as,
\begin{center}
\begin{lstlisting}
nil = text "" <+ space
space = (text " " <+ text "\n") <> nil
\end{lstlisting}
\end{center}
where \lstinline|nil| and \lstinline|space| pretty-print \lstinline|""| and
\lstinline|" "| respectively, but represent zero-or-more and one-or-more
whitespaces in parsing. FliPpr also allows annotations for optional
parentheses in programs. These features enable a user to easily define a
pretty-printer and a parser that also recognizes ``ugly'' strings; after a
round trip through the parser and printer, such 
strings will be mapped to a canonical representative modulo whitespace and parentheses quotienting.

@@

\begin{acks}
  We thank the anonymous reviewers for their useful feedback and discussions and Nate Foster for extensive assistance in integrating
  QRE-enhanced \QOpt{} into Boomerang.  This research has been supported in part by DARPA
  award FA8750-17-2-0028 and ONR 568751 (SynCrypt).
\end{acks}
